

library(tidyverse)    
library(knitr)        
library(rmarkdown)    
library(glue)         
library(scales)       

eurovision <- read_csv("data/eurovision.csv")

create_rmd_template <- function() {
  template_content <- '---
title: "{{{country_name}}} - Eurovision Profile"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
params:
  country_name: "Ukraine"
  years_participated: 0
  years_qualified: 0
  years_won: 0 
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(scales)
```

```{r calculate_metrics, include=FALSE}
years_participated <- function(country) {
  eurovision %>% filter(to_country == country) %>% 
    summarise(years_participated = n_distinct(year))
} 

  years_qualified <- function(country) {
  eurovision %>% filter(to_country == country) %>% 
    filter(!is.na(place_final)) %>% 
  summarise(years_qualified = n_distinct(year))
} 

  years_won <- function(country) {
  eurovision %>% filter(to_country == country) %>% 
  filter(place_final == 1) %>% 
  summarise(years_won = n_distinct(year))
}

  graph_final_place_distribution <- function(country) {
  eurovision %>% filter(to_country == country) %>% 
  filter(!is.na(place_final)) %>% 
  group_by(year) %>% 
  slice(1) %>% 
  ungroup() %>% 
  count(place_final) %>%   
  ggplot(aes(x = place_final, y = n)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n),
    vjust = 0, 
    nudge_y = 0.1,
    color = "black") +
  scale_x_continuous(breaks = seq(1, 27, by = 1)) +
  labs(title = paste0("Final place distribution for ", country), 
       x = "Place in the Grand Final", 
       y = "Number of times place acquired") +
  theme_classic()
} 
  
  best_performance <- function(country) {
  eurovision %>% group_by(year, round) %>% 
  mutate(total_possible = sum(total_points)) %>% 
  filter(to_country == country) %>% 
  mutate(n_points = sum(total_points)) %>% 
  slice(1) %>% 
  summarise(pct_points = n_points*100/total_possible) %>% 
  arrange(desc(pct_points)) %>% 
  head(1) %>% 
  select(round, year) %>% 
    unite(col = round_year, year, round, sep = " ") %>% 
    pull(round_year)
}
  
  graph_loyal_besties <- function(country) {
  eurovision %>% filter(to_country == country) %>% 
  group_by(from_country) %>% 
  summarise(min_points = min(total_points)) %>% 
  filter(min_points != 0) %>% 
  arrange(desc(min_points)) %>% 
  ggplot(aes(x = reorder(from_country, -min_points), y = min_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(aes(label = min_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Minimum points from countries that have never given ", country, " zero points"), 
       x = "Countries", 
       y = "Minimum points") +
  theme_classic()
}


```


## Overview
This is a Eurovision Song Contest Profile for **{{{params$country_name}}}** based on 1957-2024 data.

## Participation 
**{{{params$country_name}}}** has participated in Eurovision **{{{params$years_participated}}}** times and competed in the final round **{{{params$years_qualified}}}** times. 

## Successes
**{{{params$country_name}}}** has won **{{{params$years_won}}}** times. Its most successful performance took place in the **{{{params$best_performance}}}**.

## Final place distribution 
Below is a chart showing how many times **{{{params$country_name}}}** earned a certain place in the final round of Eurovision:
```{r final_place_distribution, fig.width=10, fig.height=6}
graph_final_place_distribution(params$country_name)
```

## Loyal besties
Below are the countries that have NEVER given **{{{params$country_name}}}** zero points. The chart shows the minimum number of points those countries gave, instead. 
```{r loyal_besties, fig.width=10, fig.height=6}
graph_loyal_besties(params$country_name)
```


---

'
  
  dir.create("templates", showWarnings = FALSE)
  
  writeLines(template_content, "templates/country_template.Rmd")
  
  return("templates/country_template.Rmd")
}

# Function to generate a report for a single country
generate_country_report <- function(country_name, data) {
  template_path <- "templates/country_template.Rmd"
  if(!file.exists(template_path)) {
    template_path <- create_rmd_template()
  }
  
  
  
  
  
  # Filter data for the specific receiving country
  
  years_participated <- function(country) {
  eurovision %>% filter(to_country == country) %>% 
    summarise(years_participated = n_distinct(year))
} 

  years_qualified <- function(country) {
  eurovision %>% filter(to_country == country) %>% 
    filter(!is.na(place_final)) %>% 
  summarise(years_qualified = n_distinct(year))
} 

  years_won <- function(country) {
  eurovision %>% filter(to_country == country) %>% 
  filter(place_final == 1) %>% 
  summarise(years_won = n_distinct(year))
}

  graph_final_place_distribution <- function(country) {
  eurovision %>% filter(to_country == country) %>% 
  filter(!is.na(place_final)) %>% 
  group_by(year) %>% 
  slice(1) %>% 
  ungroup() %>% 
  count(place_final) %>%   
  ggplot(aes(x = place_final, y = n)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n),
    vjust = 0, 
    nudge_y = 0.1,
    color = "black") +
  scale_x_continuous(breaks = seq(1, 27, by = 1)) +
  labs(title = paste0("Final place distribution for ", country), 
       x = "Place in the Grand Final", 
       y = "Number of times place acquired") +
  theme_classic()
} 
  
  best_performance <- function(country) {
  eurovision %>% group_by(year, round) %>% 
  mutate(total_possible = sum(total_points)) %>% 
  filter(to_country == country) %>% 
  mutate(n_points = sum(total_points)) %>% 
  slice(1) %>% 
  summarise(pct_points = n_points*100/total_possible) %>% 
  arrange(desc(pct_points)) %>% 
  head(1) %>% 
  select(round, year) %>% 
    unite(col = round_year, year, round, sep = " ") %>% 
    pull(round_year)
}
  
  graph_loyal_besties <- function(country) {
  eurovision %>% filter(to_country == country) %>% 
  group_by(from_country) %>% 
  summarise(min_points = min(total_points)) %>% 
  filter(min_points != 0) %>% 
  arrange(desc(min_points)) %>% 
  ggplot(aes(x = reorder(from_country, -min_points), y = min_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(aes(label = min_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Minimum points from countries that have never given ", country, " zero points"), 
       x = "Countries", 
       y = "Minimum points") +
  theme_classic()
}
  
  
  
  
  
  # Create the output folder if it doesn't exist
  dir.create("country_reports", showWarnings = FALSE)
  
  # Create output filename
  output_file <- paste0("country_reports/{tolower(gsub(' ', '_', country_name))}")
  
  # Render the report with parameters
  rmarkdown::render(
    input = template_path,
    output_file = paste0(output_file, ".html"),
    params = list(
      country_name = country_name,
      country_data = country_data
    ),
    envir = new.env()
  )
  
  return(output_file)
}
  
  return(output_file)
}

# Generate reports for all countries
generate_all_reports <- function(data) {
  # Get all unique receiving countries
  all_countries <- unique(data$to_country)
  
  # First, create the template
  create_rmd_template()
  
  # Generate a report for each country
  map_chr(all_countries, ~generate_country_report(.x, data))
}

# Execute the function to generate all reports
# Uncomment the next line when ready to run
# generated_files <- generate_all_reports(eurovision_data)

# Print the generated files
# print(generated_files)

# If you want to process just one country for testing:
# test_file <- generate_country_report("Sweden", eurovision_data)
# print(test_file)
  
  # Write markdown to file
  writeLines(md_content, md_filename)
  
  # Render to HTML (optional)
  rmarkdown::render(md_filename, output_format = "html_document", 
                    output_dir = "country_reports")
  
  return(md_filename)
}

# Generate reports for all countries
generate_all_reports <- function(data) {
  # Get all unique receiving countries
  all_countries <- unique(data$to_country)
  
  # Generate a report for each country
  map_chr(all_countries, ~generate_country_report(.x, data))
}

# Execute the function to generate all reports
# Uncomment the next line when ready to run
# generated_files <- generate_all_reports(eurovision_data)

# Print the generated files
# print(generated_files)

# If you want to process just one country for testing:
test_file <- generate_country_report("Sweden", eurovision_data)
print(test_file)