---
title: "Eurovision Song Contest country profile"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
date: "2025-05-02"
params:
  country_name: "Andorra" 
---

```{r setup, include = FALSE}
library(tidyverse)
library(rmarkdown)
library(ggplot2)
library(patchwork)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r datasets, include = FALSE}
eurovision <- read_csv("data/eurovision.csv")
```

```{r countries, include = FALSE}
countries <- unique(eurovision$to_country)
```

```{r dataset_two, include = FALSE}
country_to_data <- eurovision %>%
  filter(to_country == params$country_name)
```

```{r dataset_three, include = FALSE}
country_from_data <- eurovision %>%
  filter(from_country == params$country_name)
```

```{r years_participated, include = FALSE}
years_participated <- country_to_data %>%
  summarise(years_participated = n_distinct(year))
```

```{r years_qualified, include = FALSE}
years_qualified <- country_to_data %>% 
  filter(!is.na(place_final)) %>% 
  summarise(years_qualified = n_distinct(year))
```

```{r qualification_rate, include = FALSE}
qualification_rate <- country_to_data %>% 
  summarise(years_participated = n_distinct(year),
            years_qualified = n_distinct(year[!is.na(place_final)])) %>%
  mutate(qualification_rate = years_qualified * 100 / years_participated) %>%
  head(1) %>% 
  select(qualification_rate) %>% 
  mutate(qualification_rate = round(qualification_rate, 1))
```

```{r years_won, include = FALSE}
years_won <- country_to_data %>% 
  filter(place_final == 1) %>% 
  summarise(years_won = n_distinct(year))
```

```{r best_performance, include = FALSE}
best_performance <- eurovision %>% group_by(year, round) %>% 
  mutate(total_possible = sum(total_points)) %>% 
  filter(to_country == params$country_name) %>% 
  mutate(n_points = sum(total_points)) %>% 
  slice(1) %>% 
  summarise(pct_points = n_points*100/total_possible) %>% 
  arrange(desc(pct_points)) %>% 
  head(1) %>% 
  unite(col = round_year, year, round, sep = " ") %>% 
  mutate(pct_points = round(pct_points, 1))
```

# `r params$country_name`


## Participation and successes

`r params$country_name` has participated in the Eurovision Song Contest `r years_participated$years_participated` times. It made it to the final round `r years_qualified$years_qualified` times - that's `r qualification_rate$qualification_rate`%. 

`r params$country_name` has won `r years_won$years_won` times. Its most successful performance took place in the `r best_performance$round_year`, when the country got `r best_performance$pct_points`% of all possible points. 

## Final place distribution 

Here's how many times `r params$country_name` earned a particular place in the final round of Eurovision: 

```{r graph_final_place_distribution, fig.width=7, fig.height=4, echo = FALSE}
graph_final_place_distribution <- country_to_data %>% 
  filter(!is.na(place_final)) %>% 
  group_by(year) %>% 
  slice(1) %>% 
  ungroup() %>% 
  count(place_final)

if (nrow(graph_final_place_distribution) > 0) {
  ggplot(graph_final_place_distribution, aes(x = place_final, y = n)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n),
    vjust = 0, 
    nudge_y = 0.1,
    color = "black") +
  scale_x_continuous(breaks = seq(1, 27, by = 1)) +
  labs(title = paste0("Final place distribution for ", params$country_name), 
       x = "Place in the Grand Final", 
       y = "Number of times place acquired") +
  theme_classic()
} else {
  cat("This country has not participated in the final round.")
}
```

## Loyal besties 

Below are the countries that have NEVER given `r params$country_name` zero points. The bar chart shows the lowest score those countries gave `r params$country_name` instead.

```{r graph_loyal_besties, fig.width=7, fig.height=4, echo = FALSE}
graph_loyal_besties <- country_to_data %>% 
  group_by(from_country) %>% 
  summarise(min_points = min(total_points)) %>% 
  filter(min_points != 0) %>% 
  arrange(desc(min_points))

if (nrow(graph_loyal_besties) > 0) {
  ggplot(graph_loyal_besties, aes(x = reorder(from_country, -min_points), 
                                y = min_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(aes(label = min_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Minimum points from countries that have never given ", params$country_name, " zero points"), 
       x = "Countries", 
       y = "Minimum points") +
  theme_classic()
} else {
  cat(paste0("All participating countries have given ", params$country_name, " zero points at least once."))
}
```

## Only zeros 

On the other hand, here are the countries that have never given `r params$country_name` more than zero points. 

```{r only_zeros, echo = FALSE}
only_zeros <- country_to_data %>% 
  group_by(from_country) %>% 
  summarise(n_total = sum(total_points)) %>% 
  arrange(n_total) %>% 
  filter(from_country != params$country_name) %>% 
  filter(n_total == 0) %>% 
  distinct(from_country)

if (nrow(only_zeros) > 0) {
  cat("countries that have never given ", params$country_name, " more than zero points: \n\n")
  cat(paste0("- ", only_zeros$from_country, collapse = "\n"))
} else {
  cat("Just kidding! No country would do something like that to", params$country_name)
}
```

```{r graph_tele_besties, include = FALSE}
graph_tele_besties <- country_to_data %>% 
  filter(year >=2016) %>% 
  group_by(from_country) %>% 
  summarise(n_tele_points = sum(tele_points)) %>% 
  arrange(desc(n_tele_points)) %>% 
  head(5) 

if (nrow(graph_tele_besties) > 0) {
  ggplot(graph_tele_besties, aes(x = reorder(from_country, -n_tele_points), 
                                 y = n_tele_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_tele_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Top five televoter friends of ", params$country_name), 
       x = "Televoting friend", 
       y = "Points") +
  theme_classic()
} else {
  ggplot() + theme_void() + labs(title = "This country has not participated since 2016 when televotes became public.")
}

p1 <- if (nrow(graph_tele_besties) > 0) {
  ggplot(graph_tele_besties, aes(x = reorder(from_country, -n_tele_points), 
                                 y = n_tele_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_tele_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Top five televoter friends of ", params$country_name), 
       x = "Televoting friend", 
       y = "Points") +
  theme_classic()
} else {
  ggplot() + theme_void() + labs(title = "This country has not participated since 2016 when televotes became public.")
}
```

```{r graph_tele_topoints, include = FALSE}
graph_tele_topoints <- country_from_data %>% 
  filter(year >=2016) %>%
  group_by(to_country) %>% 
  summarise(n_tele_points = sum(tele_points)) %>% 
  arrange(desc(n_tele_points)) %>% 
  head(5)
  
if (nrow(graph_tele_topoints) > 0) {
  ggplot(graph_tele_topoints, aes(x = reorder(to_country, -n_tele_points), 
                                      y = n_tele_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_tele_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0(params$country_name, " televoters gave points to"), 
       x = "Televoting friend", 
       y = "Points") +
  theme_classic()
} else {
  ggplot() + theme_void() + labs(title = "This country has not participated since 2016 when televotes became public.")
} 

p2 <- if (nrow(graph_tele_topoints) > 0) {
  ggplot(graph_tele_topoints, aes(x = reorder(to_country, -n_tele_points), 
                                      y = n_tele_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_tele_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0(params$country_name, " televoters gave points to"), 
       x = "Televoting friend", 
       y = "Points") +
  theme_classic()
} else {
  geom_blank()
} 
```

```{r graph_jury_besties, include = FALSE}
graph_jury_besties <- country_to_data %>% 
  filter(year >=2016) %>% 
  mutate(jury_points = replace_na(jury_points, 0)) %>%
  group_by(from_country) %>% 
  summarise(n_jury_points = sum(jury_points)) %>% 
  arrange(desc(n_jury_points)) %>% 
  head(5)
  
if (nrow(graph_jury_besties) > 0) {
  ggplot(graph_jury_besties, aes(x = reorder(from_country, -n_jury_points), 
                                 y = n_jury_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_jury_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Top five jury friends of ", params$country_name), 
       x = "Jury friend", 
       y = "Points") +
  theme_classic()
} else {
  ggplot() + theme_void() + labs(title = "This country has not participated since 2016 when jury votes became a separate category.")
} 

p3 <- if (nrow(graph_jury_besties) > 0) {
  ggplot(graph_jury_besties, aes(x = reorder(from_country, -n_jury_points), 
                                 y = n_jury_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_jury_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Top five jury friends of ", params$country_name), 
       x = "Jury friend", 
       y = "Points") +
  theme_classic()
} else {
  ggplot() + theme_void() + labs(title = "This country has not participated since 2016 when jury votes became a separate category.")
} 
```

```{r graph_jury_topoints, include = FALSE}
graph_jury_topoints <- country_from_data %>% 
  filter(year >=2016) %>% 
  mutate(jury_points = replace_na(jury_points, 0)) %>% 
  group_by(to_country) %>% 
  summarise(n_jury_points = sum(jury_points)) %>% 
  arrange(desc(n_jury_points)) %>% 
  head(5) 

if(nrow(graph_jury_topoints) > 0) {
  ggplot(graph_jury_topoints, aes(x = reorder(to_country, -n_jury_points), 
                                  y = n_jury_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_jury_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Countries ", params$country_name, " jury gave points to"), 
       x = "Jury friend", 
       y = "Points") +
  theme_classic()
} else {
  ggplot() + theme_void() + labs(title = "This country has not participated since 2016 when jury votes became a separate category.")
}

p4 <- if(nrow(graph_jury_topoints) > 0) {
  ggplot(graph_jury_topoints, aes(x = reorder(to_country, -n_jury_points), 
                                  y = n_jury_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_jury_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Countries ", params$country_name, " jury gave points to"), 
       x = "Jury friend", 
       y = "Points") +
  theme_classic()
} else {
  geom_blank()
}  
```

```{r graph_overall_besties, include = FALSE}
graph_overall_besties <- country_to_data %>% 
  group_by(from_country) %>% 
  summarise(n_points = sum(total_points)) %>% 
  arrange(desc(n_points)) %>% 
  head(5) 

ggplot(graph_overall_besties, aes(x = reorder(from_country, -n_points), 
                                  y = n_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Top five friends of ", params$country_name), 
       x = "Countries", 
       y = "Points") +
  theme_classic()

p5 <- ggplot(graph_overall_besties, aes(x = reorder(from_country, -n_points), 
                                  y = n_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Top five friends of ", params$country_name), 
       x = "Countries", 
       y = "Points") +
  theme_classic()
```

```{r graph_overall_topoints, include = FALSE}
graph_overall_topoints <- country_from_data %>% 
  group_by(to_country) %>% 
  summarise(n_points = sum(total_points)) %>% 
  arrange(desc(n_points)) %>% 
  head(5)

ggplot(graph_overall_topoints, aes(x = reorder(to_country, -n_points), 
                                   y = n_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Countries ", params$country_name, " gave points to"), 
       x = "Countries", 
       y = "Points") +
  theme_classic()
  
p6 <- ggplot(graph_overall_topoints, aes(x = reorder(to_country, -n_points), 
                                   y = n_points)) +
  geom_col(fill = "darkseagreen3") +
  geom_text(
    aes(label = n_points),
    position = position_stack(vjust = 0.5),
    color = "black") +
  labs(title = paste0("Countries ", params$country_name, " gave points to"), 
       x = "Countries", 
       y = "Points") +
  theme_classic()
```

## Birds of a feather sing together

The European Broadcasting Union started separating televote (votes cast by audience) and jury vote (votes cast by official representatives of each participating country) in 2016. 
Below is a chart that shows top five countries whose public voted for `r params$country_name` and top five countries that `r params$country_name` public voted for the most between 2016 and 2024:

```{r patchwork_one, echo = FALSE}
p1 + p2
```

Below is a chart that shows top five countries whose jury voted for `r params$country_name` and top five countries that `r params$country_name` jury voted for the most between 2016 and 2024:

```{r patchwork_two, echo = FALSE}
p3 + p4
```

Finally, here is a chart that shows overall (televote and jury vote combined) top five countries that voted for `r params$country_name` and top five countries that `r params$country_name` voted for the most between 1957 and 2024:

```{r patchwork_three, echo = FALSE}
p5 + p6
```

```{r render_all_countries, eval=FALSE}

# Now, I am going to try and render this for all of my countries. I asked Claude how to do that and it came up with this 

dir.create("country_reports", showWarnings = FALSE)

for(country in countries) {
  rmarkdown::render(
    "scripts/04_actual_eurovision_markdown.Rmd",  
    output_file = file.path("country_reports", paste0(country, "_eurovision_profile.html")),
    params = list(country_name = country)
  )
  cat("Rendered report for", country, "\n")
}

#This did not work, so I am turning to Claude again

cat("Current working directory:", getwd(), "\n")
cat("Does 'country_reports' exist in current directory?", dir.exists("country_reports"), "\n")
cat("List of files/folders in current directory:", "\n")
print(list.files())

# Create country_reports directory if it doesn't exist
dir.create("country_reports", showWarnings = FALSE)
cat("Directory created or already exists. Checking again:", dir.exists("country_reports"), "\n")

# Now let's create a simple test file to see if we can write to this directory
test_file <- file.path("country_reports", "test.txt")
write.table("test", file = test_file)
cat("Test file created at:", normalizePath(test_file), "\n")

eurovision <- read.csv("data/eurovision.csv")  

countries <- unique(eurovision$to_country)  

# Get absolute path to Rmd file
rmd_file <- normalizePath("scripts/04_actual_eurovision_markdown.Rmd", mustWork = TRUE)
cat("Using Rmd file:", rmd_file, "\n")

# Create output directory (absolute path)
output_dir <- file.path(getwd(), "country_reports")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
cat("Output directory:", normalizePath(output_dir), "\n")

# Loop through countries - use only a few for testing
test_countries <- head(countries, 3)  # Just do 3 for testing
for(country in test_countries) {
  output_file <- file.path(output_dir, paste0(country, "_eurovision_profile.html"))
  
  cat("Rendering for country:", country, "\n")
  cat("Output will be:", output_file, "\n")
  
  tryCatch({
    rmarkdown::render(
      input = rmd_file,
      output_file = output_file,
      params = list(country_name = country)
    )
    cat("Successfully rendered for", country, "\n")
  }, error = function(e) {
    cat("Error rendering for", country, ":", e$message, "\n")
  })
}

# I ran this and got this error message: "Error rendering for Denmark : 'data/eurovision.csv' does not exist in current working directory ('C:/Users/marii/Documents/GitHub/ADJ_final/scripts')."

# And now, I understand what the problem is! R was looking for my data in the scripts folder, but my data is in the data folder. I am going back to Claude on how to fix this.

# Check current working directory
project_dir <- getwd()
cat("Current working directory:", project_dir, "\n")

# Load your dataset from the correct location
eurovision <- read.csv(file.path(project_dir, "data/eurovision.csv"))
cat("Dataset loaded successfully\n")

# Get unique country names
countries <- unique(eurovision$to_country)  

# Paths for input and output
rmd_file <- file.path(project_dir, "scripts/04_actual_eurovision_markdown.Rmd")
output_dir <- file.path(project_dir, "country_reports")
dir.create(output_dir, showWarnings = FALSE)

# Loop through countries
for(country in countries) {
  output_file <- paste0(country, "_eurovision_profile.html")
  
  cat("Rendering for country:", country, "\n")
  
  rmarkdown::render(
    input = rmd_file,
    output_file = file.path(output_dir, output_file),
    params = list(country_name = country),
    knit_root_dir = project_dir  
  )
  cat("Successfully rendered for", country, "\n")
}

#Yayyyy! It worked! I can't believe it actually happened!!
# This is all I have intended to do for this final project. You can now find Eurovision information for any of the 52 countries that have ever participated. Those reports are in the country_reports folder. Once Eurovision 2025 happens, I will definitely add in the new data and rerun the code to make it up to date!
#I apologize for the amount I have made you read, but I hope you found the ride as fascinating as I did.
```
